Identification Paper

### 2023-09-02 my_paper/Hadah_2023.tex
first commit

\documentclass[12pt,english]{article}

\input{fragments/preamble_packages.tex}
\input{fragments/preamble_layout.tex}
\input{fragments/preamble_citation.tex}
\bibliography{bib/AttitudesPaper}

\makeatother

@@

\title{The Effect of Racial and Ethnic Attitudes on Hispanic Identity in the U.S}

\author{\href{https://hussainhadah.com/}{Hussain Hadah} \thanks{Hussain Hadah: Department of Economics, Tulane University, 6823 St. Charles Ave., New Orleans, LA 70118, United States (e-mail: \href{mailto:hhadah@tulane.edu}{hhadah@tulane.edu}). I thank Professors Willa Friedman, Chinhui Juhn, Vikram Maheshri, and Yona Rubinstein for their support and advice. I also thank Aimee Chin, Steven Craig, German Cubas, Elaine Liu, Fan Wang, and the participants of the Applied Microeconomics Workshop at the University of Houston, and European Society for Population Economics (ESPE) for helpful feedback.}}

\maketitle

\begin{center}
\href{https://hhadah.github.io/Attitudes-and-Identity/my_paper/Hadah-Attitudes.pdf}{\textcolor{green!15!black!30!blue}{\footnotesize{\textsc{Click Here to Get the Most Updated Version}}}}
\end{center}

\begin{abstract}
I study the determinants of the choice to identify as Hispanic among those who could—those whose parents, grandparents, or selves were born in a Spanish-speaking country. I find that individuals with Hispanic ancestry are significantly less likely to self-identify as Hispanic if they live in states with high levels of implicit ethnic bias. A one standard deviation increase in bias decreases self-reported Hispanic identity by 7 and 13 percentage points for first and second-generation Hispanics, respectively. These effects are more prominent among second-generation immigrants with both parents born in a Spanish-speaking country than among children of inter-ethnic parents. These findings have implications for the interpretation of research on ethnic gaps in economic outcomes and the correct counting of the population.

\noindent\textbf{Keywords}: Economics of Minorities, Race, and Immigrants; Discrimination and Prejudice

\noindent\textbf{JEL Classification}: I310, J15, J71

@@

\vfill
\pagebreak{}

@@

The self-reported Hispanic population in the United States has tripled over the last three decades as Hispanics overtook Blacks as the largest minority group in America.\footnote{The 2020 Census counted more than 62 million Hispanics---19 percent of the population---triple the number of Hispanics counted three decades earlier \autocite{floodsarahIntegratedPublicUse2021a}. The Hispanic population numbers are based on the author's calculations from the Current Population Survey and US Census data.} An extensive literature on Hispanics--non-Hispanic White gaps has emerged, finding differences in health \autocite{antecolUnhealthyAssimilationWhy2006,antmanEthnicAttritionObserved2016}, wages \autocite{trejoWhyMexicanAmericans1997}, and schooling \autocite{antmanHispanicAmericansLabor2022}. However, defining and measuring these racial and ethnic groups is not straightforward, especially when considering self-reported identity. To the extent that self-report Hispanic identity is negatively selected, these gaps could be biased upward.

@@

To explore how individual characteristics and social attitudes toward Hispanics affect self-reported Hispanic identity. I use identity and ancestry information from the Current Population Survey (CPS) along with a proxy for state-level bias from Harvard's Project Implicit Association Test (IAT). \footnote{The IAT data is retrieved from Harvard's Project Implicit \autocite{greenwaldMeasuringIndividualDifferences1998}. The implicit bias toward minorities, as measured by IAT, is widely used by psychologists and is growing in use among economists. IAT scores were shown to be correlated with economic outcomes \autocite{chettyRaceEconomicOpportunity2020,gloverDiscriminationSelfFulfillingProphecy2017}, voting behavior \autocite{friesePredictingVotingBehavior2007}, and health \autocite{leitnerRacialBiasAssociated2016}.} I motivate my analysis with a simple model in the vein of \textcite{akerlofEconomicsIdentity2000}. This makes explicit a path through which actions affect individuals' utility via their identity and introduces an externality where the actions of others---or prejudice---have different effects on a person's well-being and identity. Therefore, if a person can choose their identity credibly and this choice is affected by the prejudice of others, then they will choose it to maximize their outcomes.

@@

This paper fits in the economics of immigration and assimilation. \textcite{abramitzkyCulturalAssimilationAge2016} measured the speed at which immigrants from Europe, Asia, and Latin America assimilate in the United States. They find that assimilation increases over time.\footnote{For more on immigrant assimilation, see \textcite{abramitzkyLeavingEnclaveHistorical2020,abramitzkyIntergenerationalMobilityImmigrants2019,abramitzkyDiscriminationReturnsCultural2020,abramitzkyNationImmigrantsAssimilation2014}} \textcite{foukaImmigrantsAmericansRace2022} investigated the effect of the inflow of Black Americans migrating from the South to the North on the assimilation of European immigrants. The authors found that immigrants in places that received more Black migrants assimilated faster.  \textcite{mengIntermarriageEconomicAssimilation2005} studied the effect of intermarriage on assimilation and found that immigrants who intermarry earn significantly more than those in an endogamous marriage. \textcite{antmanEthnicAttritionObserved2016} show that among immigrants from Mexico, the least economically successful self-identify as being of Mexican origin, while the most successful do not.

Other researchers studied the assimilation of Hispanic immigrants. \textcite{antecolUnhealthyAssimilationWhy2006} documented an interesting puzzle where non-native-born Hispanics have better health outcomes than native-born Hispanics, and \textcite{trejoWhyMexicanAmericans1997} showed that Mexican men earn substantially less than Whites.\footnote{The Hispanic health paradox has led many researchers to try to explain it \autocite{giuntellaAssimilationHealthEvidence2016,giuntellaAccelerationImmigrantUnhealthy2017,giuntellaReasonImmigrationImmigrants2018,giuntellaWhyDoesHealth2017a,antmanEthnicAttritionObserved2016,antmanEthnicAttritionAssimilation2020}.} \textcite{smithAssimilationLatinoGenerations2003} offered a more optimistic view of the assimilation of Hispanic immigrants. The longer Hispanic and Latino immigrants reside in the US, the more they can close the educational gap with White men. Moreover, some of the poor showings of how well Hispanic immigrants assimilate in the United States could be explained by ethnic attrition and the use of self-reported Hispanic identity to study Hispanics \autocite{duncanComplexityImmigrantGenerations2017,duncanWhoRemainsMexican2011,mengIntermarriageEconomicAssimilation2005,duncanIdentifyingLaterGenerationDescendants2018,duncanSocioeconomicIntegrationImmigrant2018,antmanEthnicAttritionObserved2016,antmanEthnicAttritionAssimilation2020}. The ethnic attrition was driven by the children of inter-ethnic marriages \autocite{mengIntermarriageEconomicAssimilation2005,duncanEthnicIdentificationIntermarriage2005}. Once the attrition was accounted for, Hispanic immigrants would appear healthier and thus more assimilated than previously thought \autocite{antmanEthnicAttritionObserved2016,antmanEthnicAttritionAssimilation2020}. This research, however, does not explore whether the self-reported identity itself is a function of other factors like prejudice.

This paper is most closely related to \textcite{antmanEthnicAttritionObserved2016,antmanIncentivesIdentifyRacial2015,antmanAmericanIndianCasinos2021} where the authors studied the ethnic attrition of Hispanic immigrants and how minorities change their self-reported identity to changes in policies.\footnote{Ethnic attrition is when a person with Hispanic ancestry fails to self-identify as Hispanic.} Taking into consideration the ethnic attrition that \textcite{antmanEthnicAttritionObserved2016} document, I investigate the determinants of what drives a person to self-report, or not, their Hispanic identity. I aim to decompose some of the complexity associated with endogenous identity by exploring some of the personal and environmental determinants of identity. The empirical analysis in this paper documents how some observable, i.e., personal characteristics and societal attitudes, affect the self-reported identity of Hispanics.

@@

I discuss a conceptual framework of identity in the spirit of \textcite{akerlofEconomicsIdentity2000}. A person belongs to some ethnic group, and their actions either affirm or deny their ethnic identity. Actions that deviate from what is proscribed of the ethnic identity are costly.

@@

Where $\pmb{a_i}$ is the actions of person $i$. $\pmb{a_{-i}}$ is the actions of others that would affect $i$'s identity, i.e., societal bias. $I_i$ is the identity function. Each group has an associated set of behaviors that society proscribes them to conform to, which I denote as $\pmb{B}_{e_{i}}$.\footnote{\textcite{akerlofEconomicsIdentity2000} refer to $B_{e_{i}}$ as proscription.}

@@

In this section, I describe the datasets I use. To study the association between social attitudes and self-reported Hispanic identity, I must measure subjective and objective Hispanic identities to select a subgroup of Hispanic immigrants for analysis. Thus, I use the Integrated Public Use Microdata Series (IPUMS) Current Population Survey (CPS) \autocite{floodsarahIntegratedPublicUse2021a} and use information on ancestry to construct an objective identity measure. I also use implicit association test (IAT) as a measure of bias, in which bias will shift the costs of identifying as a particular ethnic identity (Hispanic).

@@

I measure Hispanic identity using the Current Population Survey (CPS), which allows me to construct an objective measure of the Hispanic identity of minors living with their parents. I will use the information on the place of birth, parents' place of birth, and place of birth of grandparents to construct an objective Hispanic measure.\footnote{Following the works of \textcite{antmanEthnicAttritionObserved2016,antmanEthnicAttritionAssimilation2020}.} Thus, I could perfectly identify, and construct a dataset of first-, second-, and third-generation Hispanic immigrants (see figure \ref{fig:diag} for a visual representation). This will consequently allow me to build an objective measure of the Hispanic identity of minors under the age 17 living with their parents.

@@

My analysis depends on a sub sample of the US population, I show in table (\ref{tab:hispbygen}) that I have enough observations in each generation. Consistent with the literature on ethnic attrition among Hispanics, I find significant attrition among third-generation Hispanic immigrants.\footnote{In \textcite{duncanIdentifyingLaterGenerationDescendants2018,duncanSocioeconomicIntegrationImmigrant2018, antmanEthnicAttritionObserved2016,antmanEthnicAttritionAssimilation2020}, the authors find substantial attrition among Hispanics.} These results are displayed in table (\ref{tab:hispbygen}): most first- and second-generation Hispanic immigrants self-reportedly identified as Hispanic. Of the first-generation Hispanic immigrants, 96\% self-reportedly identified as Hispanic. Similarly, 95\% of the second-generation Hispanic immigrants self-reportedly identified as Hispanic, and 85\% of third-generation Hispanic immigrants identifying as Hispanic. That is a more than threefold increase in attrition rates. Most of the attrition among third-generation Hispanics is driven by attrition among the children of inter-ethnic marriages.

@@

% first
\begin{subfigure}{.9\textwidth}
\caption{Skin Tone Implicit Association Bias Over Time}
\centering
\includegraphics[width=.9\linewidth]{figure/Bias_twostates.png} 
\label{fig:skiniat}
\end{subfigure}
% Second
\begin{subfigure}{.9\textwidth}
\caption{Self-reported Hispanic Identity Over Time}
\centering
\includegraphics[width=.9\linewidth]{figure/Bias_twostates-hisp.png} 
\label{fig:hispanic-twostates}
\end{subfigure}
\caption*{\footnotesize{These two panels show the trends in implicit bias (panel a) and self-reported Hispanic identity among Hispanic immigrants (panel b) of the least and most biased places in the data. The District of Colombia is the least biased geographical area, and North Dakota is the most biased. The bias units are in standard deviations. Self-reported Hispanic identity is among first, second, and third-generation Hispanic immigrants aged 17 and younger still living in intact families.\\
Bias data is from the 2004-2021 Harvard's Project Implicit Association Test scores. Identity data is from the 2004-2021 Current Population Survey (CPS).}}
\end{figure}
\end{center}

@@

\begin{center}
\begin{figure}[H]
\caption{Maps of State-level Implicit Association Test Bias Over Time Measure with Census Division Regional Boundaries}
\label{fig:skiniat-maps}
% first
\begin{subfigure}{.45\textwidth}
\caption{State-level Bias in 2004}
\centering
\includegraphics[width=0.9\linewidth]{figure/2004skinmap.png} 
\label{fig:skiniat-map-2004}
\end{subfigure}
\hfill%
% Second
\begin{subfigure}{.45\textwidth}
\caption{State-level Bias in 2006}
\centering
\includegraphics[width=0.9\linewidth]{figure/2006skinmap.png} 
\label{fig:skiniat-map-2006}
\end{subfigure}
\hfill%
% third
\begin{subfigure}{.45\textwidth}
\caption{State-level Bias in 2008}
\centering
\includegraphics[width=0.9\linewidth]{figure/2008skinmap.png} 
\label{fig:skiniat-map-2008}
\end{subfigure}
\hfill%
% fourth
\begin{subfigure}{.45\textwidth}
\caption{State-level Bias in 2010}
\centering
\includegraphics[width=0.9\linewidth]{figure/2010skinmap.png} 
\label{fig:skiniat-map-2010}
\end{subfigure}
\hfill%
% fifth
\begin{subfigure}{.45\textwidth}
\caption{State-level Bias in 2012}
\centering
\includegraphics[width=0.9\linewidth]{figure/2012skinmap.png} 
\label{fig:skiniat-map-2012}
\end{subfigure}
\hfill%
% sixth
\begin{subfigure}{.45\textwidth}
\caption{State-level Bias in 2014}
\centering
\includegraphics[width=0.9\linewidth]{figure/2014skinmap.png} 
\label{fig:skiniat-map-2014}
\end{subfigure}
\hfill%
% seventh
\begin{subfigure}{.45\textwidth}
\caption{State-level Bias in 2016}
\centering
\includegraphics[width=0.9\linewidth]{figure/2016skinmap.png} 
\label{fig:skiniat-map-2016}
\end{subfigure}
\hfill%
% eigth
\begin{subfigure}{.45\textwidth}
\hspace{1cm}
\caption{State-level Bias in 2018}
\centering
\includegraphics[width=0.9\linewidth]{figure/2018skinmap.png} 
\label{fig:skiniat-map-2018}
\hfill%
\end{subfigure}
\caption*{\footnotesize{In this figure, I show the state-level implicit bias in different years in the sample. Each panel presents state-level bias during a certain year. The boundaries in red represent the different Census divisions in the United States. Notice how there is a variation across states with-in a region.}}
\end{figure}
\end{center}

@@

\begin{center}
\begin{figure}[H]
\caption{Maps of State-level Implicit Association Test Bias 2004-2021 Measure with Census Division Regional Boundaries}
\includegraphics[width=\textwidth]{figure/skinmap_all.png} 
\label{fig:iat-map-all}
\caption*{\footnotesize{In this figure, I show the state-level implicit bias in the sample of IAT tests from 2004 to 2021. The boundaries in red represent the different Census divisions in the United States. Notice how there is a variation across states with-in a region.}}
\end{figure}
\end{center}

@@

I use skin tone implicit association test data to construct a measure of state-level prejudice \autocite{greenwaldMeasuringIndividualDifferences1998}. This measure has been used in the social sciences, especially in psychology. Previous work has shown that IAT test scores are hard to manipulate \autocite{egloffPredictiveValidityImplicit2002}. \footnote{Research showed that the IAT tests are correlated with economic outcomes \autocite{chettyRaceEconomicOpportunity2020,gloverDiscriminationSelfFulfillingProphecy2017}, voting behavior \autocite{friesePredictingVotingBehavior2007}, and health \autocite{leitnerRacialBiasAssociated2016}. Participation in the IAT, an online test, is voluntary. Therefore, the samples are not random and might suffer from selection bias in who decides to take the exam. However,  bias reflected by Implicit Association Test (IAT) scores as a way to measure prejudiced attitudes has been used as a proxy for prejudiced attitudes in an area \textcite{chettyRaceEconomicOpportunity2020}.} Figure (\ref{fig:skiniat}) shows a graphical representation of the bias measure over time in the most and least biased places. A lower score implies less light-skin bias, whereas a higher score implies more discrimination against dark-skinned people. One-half of a standard deviation increase in bias is equivalent to the moving from Washington DC to North Dakota in 2012. I also show the state-level average bias over time in the maps in figure (\ref{fig:skiniat-maps}) and the overall average from 2004 to 2021 in figure (\ref{fig:iat-map-all}).

@@

I am also interested in investigating the relationship between state-level bias and migration. As the CPS does not report a person's birth state, I use the 2004-2021 Censuses to construct a sample of second-generation Hispanic immigrants \autocite{floodsarahIntegratedPublicUse2021}. I construct a mover variable to indicate whether these second-generation Hispanic immigrants have moved from their birth state to another state. For this purpose, I use the following models to estimate the relationship between state-level bias and migration:

@@

\pagebreak
\begingroup
\setstretch{1.0}
%\setstretch{1.1}
\setlength\bibitemsep{0pt}
\printbibliography
\endgroup
\pagebreak

@@

@@

>>>>

\documentclass{wptemp}
\usepackage{adjustbox}
\usepackage{setspace}\doublespacing%\onehalfspacing
\usepackage{natbib}
\newcommand{\GG}[1]{}
\usepackage{booktabs}
\usepackage{siunitx}
\newcolumntype{d}{S[input-symbols = ()]}\singlespacing
\doublespacing
\usepackage{lscape}
\newcommand{\ts}{\textsuperscript}
\usepackage{threeparttable}
\usepackage{threeparttablex}

% *****************************************************************
% Estout LaTeX wrapper
% *****************************************************************

%%Original code developed by Jörg Weber: see
%% https://www.jwe.cc/2012/03/stata-latex-tables-estout/
%% and
%% https://www.jwe.cc/blog/

\let\estinput=\input % define a new input command so that we can still flatten the document

\newcommand{\estwide}[3]{
		\vspace{.75ex}{
			%\textsymbols% Note the added command here
			\begin{tabular*}
			{\textwidth}{@{\hskip\tabcolsep\extracolsep\fill}l*{#2}{#3}}
			\toprule
			\estinput{#1}
			\bottomrule
			\addlinespace[.75ex]
			\end{tabular*}
			}
		}

\newcommand{\estauto}[3]{
		\vspace{.75ex}{
			%\textsymbols% Note the added command here
			\begin{tabular}{l*{#2}{#3}}
			\toprule
			\estinput{#1}
			\bottomrule
			\addlinespace[.75ex]
			\end{tabular}
			}
		}

% Allow line breaks with \ in specialcells
\newcommand{\specialcell}[2][c]{%
    \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}
}

\newcommand{\sym}[1]{\rlap{#1}}% Thanks David Carlisle

%%%%%%%%%%% End of wrapper %%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%% TiKz %%%%%%%%%%%%%%%%%%%%%
\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows}

\tikzstyle{startstop} = [rectangle, rounded corners, minimum width=3cm, minimum height=1cm,text centered, draw=black, fill=red!30]

\tikzstyle{io} = [trapezium, trapezium left angle=70, trapezium right angle=110, minimum width=3cm, minimum height=1cm, text centered, text width =5cm, draw=black, fill=blue!30]

\tikzstyle{process} = [rectangle, minimum width=3cm, minimum height=1cm, text centered, text width =5cm, draw=black, fill=orange!30]

\tikzstyle{decision} = [diamond, minimum width=3cm, minimum height=1cm, text centered, draw=black, fill=green!30]

\tikzstyle{arrow} = [thick,->,>=stealth]

@@

\linespread{1}{\title{{\fontfamily{ptm}\selectfont{\Large{\MakeUppercase{The Effect of Racial and Ethnic Attitudes on Hispanic Identity in the U.S}}}}
\thanks{I thank Professors Willa Friedman, Chinhui Juhn, Vikram Maheshri, and Yona Rubinstein for their support and advice. I also thank Aimee Chin, Steven Craig, German Cubas, Elaine Liu, Fan Wang, and the participants of the Applied Microeconomics Workshop at the University of Houston, and European Society for Population Economics (ESPE) for helpful feedback.}
}}
	\author{{\fontfamily{pbk}\selectfont\large{\textsc{Hussain Hadah}}\thanks{Department of Economics, Tulane University, 6823 St. Charles Ave., New Orleans, LA 70118, United States, (phone: +1-602-393-8077, e-mail: \email{hhadah@tulane.edu}). Competing interests: The author declares none.}}}
	\date{\fontfamily{pbk}\selectfont\normalsize{\textsc{\today}}}
	\maketitle

\begin{abstract}

I study the determinants of the choice to identify as Hispanic among those who could—those whose parents, grandparents, or selves were born in a Spanish-speaking country. I find that individuals with Hispanic ancestry are significantly less likely to self-identify as Hispanic if they live in states with high levels of implicit ethnic bias. A one standard deviation increase in bias decreases self-reported Hispanic identity by 7 and 13 percentage points for first and second-generation Hispanics, respectively. These effects are more prominent among second-generation immigrants with both parents born in a Spanish-speaking country than among children of inter-ethnic parents. These findings have implications for the interpretation of research on ethnic gaps in economic outcomes and the correct counting of the population.

\keywords{Economics of Minorities, Race, and Immigrants; Discrimination and Prejudice.}
\JEL{I310, J15, J71}

@@

\thispagestyle{empty}
\clearpage
\pagenumbering{arabic}
\renewcommand{\thetable}{\arabic{table}}  
\renewcommand{\thefigure}{\arabic{figure}}

@@

The self-reported Hispanic population in the United States has tripled over the last three decades as Hispanics overtook Blacks as the largest minority group in America.\footnote{The 2020 Census counted more than 62 million Hispanics---19 percent of the population---triple the number of Hispanics counted three decades earlier \citep{floodsarahIntegratedPublicUse2021a}. The Hispanic population numbers are based on the author's calculations from the Current Population Survey and US Census data.} An extensive literature on Hispanics--non-Hispanic White gaps has emerged, finding differences in health \citep{antecolUnhealthyAssimilationWhy2006,antmanEthnicAttritionObserved2016}, wages \citep{trejoWhyMexicanAmericans1997}, and schooling \citep{antmanHispanicAmericansLabor2022}. However, defining and measuring these racial and ethnic groups is not straightforward, especially when considering self-reported identity. To the extent that self-report Hispanic identity is negatively selected, these gaps could be biased upward.

@@

To explore how individual characteristics and social attitudes toward Hispanics affect self-reported Hispanic identity. I use identity and ancestry information from the Current Population Survey (CPS) along with a proxy for state-level bias from Harvard's Project Implicit Association Test (IAT). \footnote{The IAT data is retrieved from Harvard's Project Implicit \citep{greenwaldMeasuringIndividualDifferences1998}. The implicit bias toward minorities, as measured by IAT, is widely used by psychologists and is growing in use among economists. IAT scores were shown to be correlated with economic outcomes \citep{chettyRaceEconomicOpportunity2020,gloverDiscriminationSelfFulfillingProphecy2017}, voting behavior \citep{friesePredictingVotingBehavior2007}, and health \citep{leitnerRacialBiasAssociated2016}.} I motivate my analysis with a simple model in the vein of \citet{akerlofEconomicsIdentity2000}. This makes explicit a path through which actions affect individuals' utility via their identity and introduces an externality where the actions of others---or prejudice---have different effects on a person's well-being and identity. Therefore, if a person can choose their identity credibly and this choice is affected by the prejudice of others, then they will choose it to maximize their outcomes.

@@

This paper fits in the economics of immigration and assimilation. \citet{abramitzkyCulturalAssimilationAge2016} measured the speed at which immigrants from Europe, Asia, and Latin America assimilate in the United States. They find that assimilation increases over time.\footnote{For more on immigrant assimilation, see \citet{abramitzkyLeavingEnclaveHistorical2020,abramitzkyIntergenerationalMobilityImmigrants2019,abramitzkyDiscriminationReturnsCultural2020,abramitzkyNationImmigrantsAssimilation2014}} \citet{foukaImmigrantsAmericansRace2022} investigated the effect of the inflow of Black Americans migrating from the South to the North on the assimilation of European immigrants. The authors found that immigrants in places that received more Black migrants assimilated faster.  \citet{mengIntermarriageEconomicAssimilation2005} studied the effect of intermarriage on assimilation and found that immigrants who intermarry earn significantly more than those in an endogamous marriage. \citet{antmanEthnicAttritionObserved2016} show that among immigrants from Mexico, the least economically successful self-identify as being of Mexican origin, while the most successful do not.

Other researchers studied the assimilation of Hispanic immigrants. \citet{antecolUnhealthyAssimilationWhy2006} documented an interesting puzzle where non-native-born Hispanics have better health outcomes than native-born Hispanics, and \citet{trejoWhyMexicanAmericans1997} showed that Mexican men earn substantially less than Whites.\footnote{The Hispanic health paradox has led many researchers to try to explain it \citep{giuntellaAssimilationHealthEvidence2016,giuntellaAccelerationImmigrantUnhealthy2017,giuntellaReasonImmigrationImmigrants2018,giuntellaWhyDoesHealth2017a,antmanEthnicAttritionObserved2016,antmanEthnicAttritionAssimilation2020}.} \citet{smithAssimilationLatinoGenerations2003} offered a more optimistic view of the assimilation of Hispanic immigrants. The longer Hispanic and Latino immigrants reside in the US, the more they can close the educational gap with White men. Moreover, some of the poor showings of how well Hispanic immigrants assimilate in the United States could be explained by ethnic attrition and the use of self-reported Hispanic identity to study Hispanics \citep{duncanComplexityImmigrantGenerations2017,duncanWhoRemainsMexican2011,mengIntermarriageEconomicAssimilation2005,duncanIdentifyingLaterGenerationDescendants2018,duncanSocioeconomicIntegrationImmigrant2018,antmanEthnicAttritionObserved2016,antmanEthnicAttritionAssimilation2020}. The ethnic attrition was driven by the children of inter-ethnic marriages \citep{mengIntermarriageEconomicAssimilation2005,duncanEthnicIdentificationIntermarriage2005}. Once the attrition was accounted for, Hispanic immigrants would appear healthier and thus more assimilated than previously thought \citep{antmanEthnicAttritionObserved2016,antmanEthnicAttritionAssimilation2020}. This research, however, does not explore whether the self-reported identity itself is a function of other factors like prejudice.

This paper is most closely related to \citet{antmanEthnicAttritionObserved2016,antmanIncentivesIdentifyRacial2015,antmanAmericanIndianCasinos2021} where the authors studied the ethnic attrition of Hispanic immigrants and how minorities change their self-reported identity to changes in policies.\footnote{Ethnic attrition is when a person with Hispanic ancestry fails to self-identify as Hispanic.} Taking into consideration the ethnic attrition that \citet{antmanEthnicAttritionObserved2016} document, I investigate the determinants of what drives a person to self-report, or not, their Hispanic identity. I aim to decompose some of the complexity associated with endogenous identity by exploring some of the personal and environmental determinants of identity. The empirical analysis in this paper documents how some observable, i.e., personal characteristics and societal attitudes, affect the self-reported identity of Hispanics.

@@

I discuss a conceptual framework of identity in the spirit of \citet{akerlofEconomicsIdentity2000}. A person belongs to some ethnic group, and their actions either affirm or deny their ethnic identity. Actions that deviate from what is proscribed of the ethnic identity are costly.

@@

Where $\pmb{a_i}$ is the actions of person $i$. $\pmb{a_{-i}}$ is the actions of others that would affect $i$'s identity, i.e., societal bias. $I_i$ is the identity function. Each group has an associated set of behaviors that society proscribes them to conform to, which I denote as $\pmb{B}_{e_{i}}$.\footnote{\citet{akerlofEconomicsIdentity2000} refer to $B_{e_{i}}$ as proscription.}

@@

In this section, I describe the datasets I use. To study the association between social attitudes and self-reported Hispanic identity, I must measure subjective and objective Hispanic identities to select a subgroup of Hispanic immigrants for analysis. Thus, I use the Integrated Public Use Microdata Series (IPUMS) Current Population Survey (CPS) \citep{floodsarahIntegratedPublicUse2021a} and use information on ancestry to construct an objective identity measure. I also use implicit association test (IAT) as a measure of bias, in which bias will shift the costs of identifying as a particular ethnic identity (Hispanic).

@@

I measure Hispanic identity using the Current Population Survey (CPS), which allows me to construct an objective measure of the Hispanic identity of minors living with their parents. I will use the information on the place of birth, parents' place of birth, and place of birth of grandparents to construct an objective Hispanic measure.\footnote{Following the works of \citet{antmanEthnicAttritionObserved2016,antmanEthnicAttritionAssimilation2020}.} Thus, I could perfectly identify, and construct a dataset of first-, second-, and third-generation Hispanic immigrants (see figure \ref{fig:diag} for a visual representation). This will consequently allow me to build an objective measure of the Hispanic identity of minors under the age 17 living with their parents.

@@

My analysis depends on a sub sample of the US population, I show in table (\ref{tab:hispbygen}) that I have enough observations in each generation. Consistent with the literature on ethnic attrition among Hispanics, I find significant attrition among third-generation Hispanic immigrants.\footnote{In \citet{duncanIdentifyingLaterGenerationDescendants2018,duncanSocioeconomicIntegrationImmigrant2018, antmanEthnicAttritionObserved2016,antmanEthnicAttritionAssimilation2020}, the authors find substantial attrition among Hispanics.} These results are displayed in table (\ref{tab:hispbygen}): most first- and second-generation Hispanic immigrants self-reportedly identified as Hispanic. Of the first-generation Hispanic immigrants, 96\% self-reportedly identified as Hispanic. Similarly, 95\% of the second-generation Hispanic immigrants self-reportedly identified as Hispanic, and 85\% of third-generation Hispanic immigrants identifying as Hispanic. That is a more than threefold increase in attrition rates. Most of the attrition among third-generation Hispanics is driven by attrition among the children of inter-ethnic marriages.

@@

% first
\begin{subfigure}{.9\textwidth}
\caption{Skin Tone Implicit Association Bias Over Time}
\centering
\includegraphics[width=.9\linewidth]{figure/Bias_twostates.png} 
\label{fig:skiniat}
\end{subfigure}
% Second
\begin{subfigure}{.9\textwidth}
\caption{Self-reported Hispanic Identity Over Time}
\centering
\includegraphics[width=.9\linewidth]{figure/Bias_twostates-hisp.png} 
\label{fig:hispanic-twostates}
\end{subfigure}
\flushleft\footnotesize{\note{These two panels show the trends in implicit bias (panel a) and self-reported Hispanic identity among Hispanic immigrants (panel b) of the least and most biased places in the data. The District of Colombia is the least biased geographical area, and North Dakota is the most biased. The bias units are in standard deviations. Self-reported Hispanic identity is among first, second, and third-generation Hispanic immigrants aged 17 and younger still living in intact families.}\\
\note{Bias data is from the 2004-2021 Harvard's Project Implicit Association Test scores. Identity data is from the 2004-2021 Current Population Survey (CPS).}}
\end{figure}
\end{center}

@@

\begin{center}
\begin{figure}[H]
\caption{Maps of State-level Implicit Association Test Bias Over Time Measure with Census Division Regional Boundaries}
\label{fig:skiniat-maps}
% first
\begin{subfigure}{.45\textwidth}
\caption{State-level Bias in 2004}
\centering
\includegraphics[width=0.9\linewidth]{figure/2004skinmap.png} 
\label{fig:skiniat-map-2004}
\end{subfigure}
\hfill%
% Second
\begin{subfigure}{.45\textwidth}
\caption{State-level Bias in 2006}
\centering
\includegraphics[width=0.9\linewidth]{figure/2006skinmap.png} 
\label{fig:skiniat-map-2006}
\end{subfigure}
\hfill%
% third
\begin{subfigure}{.45\textwidth}
\caption{State-level Bias in 2008}
\centering
\includegraphics[width=0.9\linewidth]{figure/2008skinmap.png} 
\label{fig:skiniat-map-2008}
\end{subfigure}
\hfill%
% fourth
\begin{subfigure}{.45\textwidth}
\caption{State-level Bias in 2010}
\centering
\includegraphics[width=0.9\linewidth]{figure/2010skinmap.png} 
\label{fig:skiniat-map-2010}
\end{subfigure}
\hfill%
% fifth
\begin{subfigure}{.45\textwidth}
\caption{State-level Bias in 2012}
\centering
\includegraphics[width=0.9\linewidth]{figure/2012skinmap.png} 
\label{fig:skiniat-map-2012}
\end{subfigure}
\hfill%
% sixth
\begin{subfigure}{.45\textwidth}
\caption{State-level Bias in 2014}
\centering
\includegraphics[width=0.9\linewidth]{figure/2014skinmap.png} 
\label{fig:skiniat-map-2014}
\end{subfigure}
\hfill%
% seventh
\begin{subfigure}{.45\textwidth}
\caption{State-level Bias in 2016}
\centering
\includegraphics[width=0.9\linewidth]{figure/2016skinmap.png} 
\label{fig:skiniat-map-2016}
\end{subfigure}
\hfill%
% eigth
\begin{subfigure}{.45\textwidth}
\hspace{1cm}
\caption{State-level Bias in 2018}
\centering
\includegraphics[width=0.9\linewidth]{figure/2018skinmap.png} 
\label{fig:skiniat-map-2018}
\hfill%
\end{subfigure}
\flushleft\footnotesize{\note{In this figure, I show the state-level implicit bias in different years in the sample. Each panel presents state-level bias during a certain year. The boundaries in red represent the different Census divisions in the United States. Notice how there is a variation across states with-in a region.}}
\end{figure}
\end{center}

@@

\begin{center}
\begin{figure}[H]
\caption{Maps of State-level Implicit Association Test Bias 2004-2021 Measure with Census Division Regional Boundaries}
\includegraphics[width=\textwidth]{figure/skinmap_all.png} 
\label{fig:iat-map-all}
\flushleft\footnotesize{\note{In this figure, I show the state-level implicit bias in the sample of IAT tests from 2004 to 2021. The boundaries in red represent the different Census divisions in the United States. Notice how there is a variation across states with-in a region.}}
\end{figure}
\end{center}

@@

I use skin tone implicit association test data to construct a measure of state-level prejudice \citep{greenwaldMeasuringIndividualDifferences1998}. This measure has been used in the social sciences, especially in psychology. Previous work has shown that IAT test scores are hard to manipulate \citep{egloffPredictiveValidityImplicit2002}. \footnote{Research showed that the IAT tests are correlated with economic outcomes \citep{chettyRaceEconomicOpportunity2020,gloverDiscriminationSelfFulfillingProphecy2017}, voting behavior \citep{friesePredictingVotingBehavior2007}, and health \citep{leitnerRacialBiasAssociated2016}. Participation in the IAT, an online test, is voluntary. Therefore, the samples are not random and might suffer from selection bias in who decides to take the exam. However,  bias reflected by Implicit Association Test (IAT) scores as a way to measure prejudiced attitudes has been used as a proxy for prejudiced attitudes in an area \citet{chettyRaceEconomicOpportunity2020}.} Figure (\ref{fig:skiniat}) shows a graphical representation of the bias measure over time in the most and least biased places. A lower score implies less light-skin bias, whereas a higher score implies more discrimination against dark-skinned people. One-half of a standard deviation increase in bias is equivalent to the moving from Washington DC to North Dakota in 2012. I also show the state-level average bias over time in the maps in figure (\ref{fig:skiniat-maps}) and the overall average from 2004 to 2021 in figure (\ref{fig:iat-map-all}).

@@

I am also interested in investigating the relationship between state-level bias and migration. As the CPS does not report a person's birth state, I use the 2004-2021 Censuses to construct a sample of second-generation Hispanic immigrants \citep{floodsarahIntegratedPublicUse2021}. I construct a mover variable to indicate whether these second-generation Hispanic immigrants have moved from their birth state to another state. For this purpose, I use the following models to estimate the relationship between state-level bias and migration:

@@

@@

% \input{tables/tab03-iat_regression_tab}

% \pagebreak
% \newpage

% \input{tables/tab04-regression_tab_bygen_skin}

% \input{tables/tab05-iat_regression_tab_byparent_type}

% \pagebreak
% \newpage

@@

\bibliography{AttitudesPaper}
